Road traffic sound level estimation from realistic urban sound mixtures by Non-negative Matrix Factorization

### 2018-02-27 paper/gloaguenAppliedNMF.tex
update missing references

Acoustic sensor networks are increasingly deployed in cities, and appear more and more as a possible tool to enrich modeled road traffic noise maps through data assimilation techniques. This, or more simply the validation of the modeled maps through measures, requires first being able to isolate from the measured sound mixtures the road traffic sound level. This task is anything but trivial because of the multiple sound sources that overlap within urban sound mixtures. In this paper, a Non-negative Matrix Factorization (NMF) framework is developed to estimate road traffic noise levels within urban sound scenes. A corpus of sound scenes is artificially built imitating real annotated recordings. The realism of the scenes is validated through a perceptual test, forming a protocol that both reproduces the sensor network outputs, and in which the actual occurrence and sound level of each source is known. Three variants of NMF are tested, namely supervised, semi-supervised, and threshold initialized NMF. While the semi-supervised approach is the most appropriated for park environments, threshold initialized NMF, which is developed especially for this research purpose, appears to be the best generic approach, allowing road traffic noise level estimation with mean average errors of 1.3 dB over the tested corpus of sound scenes.

@@

Therefore, noise measurements are increasingly used in addition to simulation to describe urban noise environments. Several measurement set-ups have been proposed in the last years, including  mobile measurements with high quality microphones \cite{manvell2004sadmam, can2014measurement}, participative sensing through dedicated smartphone applications \cite{picaut:halshs-01565214, maisonneuve2009noisetube}, or the development of fixed-sensor networks. In this latter case, the sensor networks can be based either on high-quality sensors as in \cite{mietlicki2012innovative, maijala2018environmental}, or low-cost sensors as in the DYNAMAP project \cite{dynamap_2016} or the CENSE project \cite{picaut2017characterization}. The costs and benefits of each protocol are discussed. Mobile and participatory measures increase spatial coverage at low cost, but lack temporal representativeness. Fixed networks are very reliable for measuring sound levels temporal variations, but allow only a small spatial coverage of the network. In addition, the low-cost sensors enable a wider deployment, but at the cost of increased uncertainties, the most extreme example being smartphone applications.

@@

All these measurement protocol enables a priori combining measures and modeling to improve the accuracy of the produced noise maps. Traffic noise maps and measurements were compared on restrictive areas in \cite{lefebvre2017traffic} and \cite{mioduszewski2011noise}. Wei et al. modify the acoustical parameters of the simulation thanks to noise measurements. Mallet et al. call for data assimilation techniques between models and measurements to reduce the uncertainty of the produced noise maps. However, these works make the implicit assumption that the noise measurements consist mainly of road traffic. In the aim to improve road traffic noise maps, the use of measurements has first to deal with the challenge to estimate correctly the road traffic sound level.
Even if road traffic is predominant on many urban areas, urban sound environments are composed of many different overlapping sound sources (passing cars, voices, footsteps, car horn, whistling birds \dots), what makes the task of estimating correctly the traffic sound level within an urban sound mixture not trivial.

@@

An other approach, followed in this paper, is to consider the Blind Source Separation paradigm, see Figure \ref{fig:source_separation}, which consists in the extraction of a specific signal inside a set of mixed signals. From the different existing methods (CASA, ICA), Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999}, seems the most relevant method for monophonic sensor networks. Dealing easily with the overlapping between the sound sources, this method approximes the magnitude spectrogram of a single signal by the product of two non negative matrices. A lot of applications can be found for musical \cite{smaragdis_non-negative_2003, benetos2006musical} and speech \cite{wilson_speech_2008, mysore2011non} contents. Dedicated to sound separation, Immani and Kasa\"i used NMF in a two steps sound separation with the help of time variant gain features. A first study \cite{} has been conducted, in which diverse NMF formula, namely the supervised, the semi-supervised, and the threshold initialized NMF, have been applied on a large set of simulated sound scenes that which mixed traffic component with specific urban sounds at calibrated sound levels. The study proved the interest of NMF for urban sound environments and compared the benefits of each approach. However, it has now to face to real urban sound scenes.

@@

\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
	\rowcolor[HTML]{C0C0C0} 
      low firm threshold $\mathbf{t_{f,2}}$ & from 0.10 to 0.45 with a 0.01 step & 36\\
   \bottomrule
\end{tabularx}

\label{tab:experimental_factorsNMF}
\end{table*}

>>>>

Acoustic sensor networks are increasingly deployed in cities, and appear more and more as a possible tool to enrich modeled road traffic noise maps through data assimilation techniques. This, or more simply the validation of the modeled maps through measures, requires first being able to isolate from the measured sound mixtures the road traffic sound level. This task is anything but trivial because of the multiple sound sources that overlap within urban sound mixtures. In this paper, a Non-negative Matrix Factorization (NMF) framework is developed to estimate road traffic noise levels within urban sound scenes. A corpus of sound scenes is artificially built imitating real annotated recordings. The realism of the scenes is validated through a perceptual test, forming a protocol that both reproduces the sensor network outputs, and in which the actual occurrence and sound level of each source is known. Three variants of NMF are tested, namely supervised, semi-supervised, and threshold initialized NMF. While the semi-supervised approach is the most appropriated for park environments, threshold initialized NMF, which is developed especially for this research purpose, appears to be the best generic approach, allowing road traffic noise level estimation with average errors of 1.3 dB over the tested corpus of sound scenes.

@@

Therefore, noise measurements are increasingly used in addition to simulation to describe urban noise environments. Several measurement set-ups have been proposed in the last years, including  mobile measurements with high quality microphones \cite{manvell2004sadmam, can2014measurement}, participative sensing through dedicated smartphone applications \cite{picaut:halshs-01565214, ventura2017evaluation}, or the development of fixed-sensor networks. In this latter case, the sensor networks can be based either on high-quality sensors as in \cite{mietlicki2012innovative, maijala2018environmental}, or low-cost sensors as in the DYNAMAP project \cite{dynamap_2016} or the CENSE project \cite{picaut2017characterization}. The costs and benefits of each protocol are discussed. Mobile and participatory measures increase spatial coverage at low cost, but lack temporal representativeness. Fixed networks are very reliable for measuring sound levels temporal variations, but allow only a small spatial coverage of the network. In addition, the low-cost sensors enable a wider deployment, but at the cost of increased uncertainties, the most extreme example being smartphone applications.

@@

All these measurement protocol enables a priori combining measures and modeling to improve the accuracy of the produced noise maps. Traffic noise maps and measurements were compared on restrictive areas in \cite{lefebvre2017traffic} and \cite{mioduszewski2011noise}. Wei et al. \cite{wei_dynamic_2016} modify the acoustical parameters of the simulation thanks to noise measurements, while  Mallet et al. \cite{ventura2017estimation} call for data assimilation techniques between models and measurements to reduce the uncertainty of the produced noise maps. However, these works make the implicit assumption that the noise measurements consist mainly of road traffic. In the aim to improve road traffic noise maps, the use of measurements has first to deal with the challenge to estimate correctly the road traffic sound level.
Even if road traffic is predominant on many urban areas, urban sound environments are composed of many different overlapping sound sources (passing cars, voices, footsteps, car horn, whistling birds \dots), what makes the task of estimating correctly the traffic sound level within an urban sound mixture not trivial.

@@

An other approach, followed in this paper, is to consider the Blind Source Separation paradigm, see Figure \ref{fig:source_separation}, which consists in the extraction of a specific signal inside a set of mixed signals. From the different existing methods (CASA, ICA), Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999}, seems the most relevant method for monophonic sensor networks. Dealing easily with the overlapping between the sound sources, this method approximes the magnitude spectrogram of a single signal by the product of two non negative matrices. A lot of applications can be found for musical \cite{smaragdis_non-negative_2003, benetos2006musical} and speech \cite{wilson_speech_2008, mysore2011non} contents. Dedicated to sound separation, Immani and Kasa\"i \cite{satoshi_innami_nmf-based_2012} used NMF in a two steps sound separation with the help of time variant gain features. A first study [article JASM] has been conducted, in which diverse NMF formula, namely the supervised, the semi-supervised, and the threshold initialized NMF, have been applied on a large set of simulated sound scenes that which mixed traffic component with specific urban sounds at calibrated sound levels. The study proved the interest of NMF for urban sound environments and compared the benefits of each approach. However, it has now to face to real urban sound scenes.

@@

\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
	\rowcolor[HTML]{C0C0C0} 
      low firm threshold $\mathbf{t_{f,2}}$ & from 0.10 to 0.45 with a 0.01 step & 36\\
   \bottomrule
\end{tabularx}
\label{tab:experimental_factorsNMF}
\end{table*}
